### 2024-04-29 sections/1-introduction.tex
careful edit

$k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets because their cardinalities are growing exponentially.
While fast algorithms for $k$-NN search on large datasets do exist, most do not exploit the geometric and topological structure inherent in these datasets.
Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.

For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
Further, there is evidence that distance functions which do not obey the triangle inequality, such as Cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.

@@

This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy Scaling Search), a set of three novel algorithms for exact $k$-NN search.
We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
We also benchmark CAKES on a genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.

@@

InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively.
The number of cells used is governed by the $n_{list}$ parameter. 
Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.

@@

This search paradigm exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \textit{metric entropy} and \textit{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search. 
Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search. We also discuss improvements to the performance of CHESS's $\rho$-NN search.

>>>>

$k$-NN search is one of the most ubiquitous classification and recommendation methods in use~\cite{fix1952discriminatory, cover1967nearest}.
Na\"{i}ve implementations of $k$-NN search, whose time complexity is linear in the dataset's cardinality, prove prohibitively slow for large datasets or those whose cardinalities are growing exponentially.
While fast algorithms for $k$-NN search on large datasets exist, most do not exploit the geometric and topological structure inherent in these datasets.
Further, such algorithms are often approximate~\cite{gao2023high}, and while approximate search may be sufficient for some applications, the need for efficient and \textit{exact} search remains~\cite{ukey2023survey}.

For example, for a majority voting classifier, approximate $k$-NN search may agree with exact $k$-NN search for large values of $k$, but may be sensitive to local perturbations for smaller values of $k$.
This is especially true when classes are not well-separated~\cite{zhang2022imbalanced}.
Further, there is evidence that distance functions which do not obey the triangle inequality, such as cosine distance, perform poorly for $k$-NN search in biomedical settings~\cite{hu2016distance};
this suggests that approximate $k$-NN search could exhibit suboptimal classification accuracy in such contexts.

@@

This paper introduces CAKES (CLAM-Accelerated $K$-NN Entropy Scaling Search), a set of three novel algorithms for exact $k$-NN search.
We also present some improvements to the clustering and $\rho$-NN search algorithms in CHESS~\cite{ishaq2019clustered}, as well as improved genericity across distance functions.
We provide a comparison of CAKES's algorithms to several state-of-the-art algorithms for similarity search, FAISS~\cite{johnson2019billion}, HNSW~\cite{malkov2016hnsw}, and ANNOY~\cite{annoy}, on datasets from the ANN-benchmarks suite~\cite{aumuller2020ann}.
We also benchmark CAKES on a large genomic dataset, the SILVA 18S dataset~\cite{10.1093/nar/gks1219}, using Levenshtein~\cite{levenshtein1966binary} distance on unaligned genomic sequences, and a radio frequency dataset, the RadioML dataset~\cite{oshea2018radioml}, using Dynamic Time Warping (DTW)~\cite{gold2018dynamic} distance on complex-valued time-series.

@@

InVerted File indexing (IVF)~\cite{faissivf, sacks1987multikey, kent1990signature} is a method for approximate $k$-NN search. 
The data are clustered into high-dimensional Voronoi cells, and whichever cell the query point falls into is then searched exhaustively, similarly to an early precursor to our work~\cite{yu2015entropy}.
The number of cells used is governed by the $n_{list}$ parameter. 
Increasing this parameter decreases the number of points being exhaustively searched, so it improves speed at the cost of accuracy.
To mitigate accuracy issues caused by a query point falling near a cell boundary, the algorithm has a tunable parameter $n_{probe}$, which specifies the number of additional adjacent or nearby cells to search.

@@

This search paradigm exploits the geometric and topological structure inherent in large datasets.
Importantly, as suggested by their name, entropy-scaling search algorithms have asymptotic complexity that scales with topological properties (such as the \textit{metric entropy} and \textit{local fractal dimension}, as defined in Section~\ref{sec:methods}) of the dataset, instead of its cardinality.
In 2019, we introduced CHESS (Clustered Hierarchical Entropy-Scaling Search)~\cite{ishaq2019clustered}, which extended entropy-scaling $\rho$-NN search from a flat clustering approach to a tree-based hierarchical clustering approach.
CLAM (Clustering, Learning and Approximation with Manifolds), originally developed to allow ``manifold mapping'' for anomaly detection~\cite{ishaq2021clustered}, is a refinement of the clustering algorithm from CHESS.
In this paper, we introduce CAKES, a set of three entropy-scaling algorithms for $k$-NN search, implemented in the Rust programming language. 
Using the cluster tree constructed by CLAM, CAKES extends CHESS to perform $k$-NN search. We also discuss improvements to the performance of CHESS's $\rho$-NN search.

### 2024-04-29 sections/2-methods.tex
minor edits
more minor edits
fixed figure caption
careful edit

We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
In addition to these constraints, if the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while Cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).

The choice of an appropriate distance function varies by dataset and domain.
For example, with neural-network embeddings, one could use Euclidean (L2) or Cosine distance.
With genomic or proteomic data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.

CAKES assumes the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions, such as voids, volumes and proximity to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.

@@

We stress that this concept differs from the \textit{embedding dimension} of a dataset.
To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
However, due to physical constraints (namely, biological evolution and the chemistry of RNA replication and mutation processes), the data are constrained to a lower-dimensional manifold within this embedding space.
LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.

@@

We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
The \textit{center} is the geometric median of the points (or a smaller sample of the points) in the \textit{cluster}, and so it is a real data point.
The \textit{radius} is the maximum distance from the \textit{center} to any point in the \textit{cluster}.
Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B_X(c, r)$.

Hereafter, when we refer to the LFD of a \textit{cluster}, it is estimated at the length scale of the cluster radius and half that radius.
This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:

@@

We start by performing a divisive hierarchical clustering on the dataset using CLAM.
The procedure is similar to that outlined in CHESS, but with the following improvements:
better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).

@@

CLAM starts by building a binary tree of clusters using a divisive hierarchical clustering algorithm.

For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the \textit{center} of $C$.

@@

With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers significant improvements at search-time over the approach used in CHESS.

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.8]{images/geometry/deltas.pdf}
    \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
    \label{fig:methods:deltas}
\end{figure}

@@

\begin{figure}[ht!]
    \centering
    \includegraphics[scale=0.8]{images/geometry/overlapping-children-3.pdf}
    \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children.}
    \label{fig:methods:overlapping-children}
\end{figure}

@@

From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-life datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.

@@

We perform some na\"{i}ve auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
We start by taking the center of every cluster at a low depth, (e.g. 10), in the cluster tree as a query.
This gives us a small, representative sample of the dataset.
Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{sec:methods:knn-search}.
We select the fastest algorithm over all the queries as the optimal algorithm for that dataset and value of $k$.
Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.

>>>>

We define a \textit{distance function} $f : \textbf{X} \times \textbf{X} \mapsto \mathbb{R}^+ \ \cup \ \{0\}$ which, given two points, deterministically returns a finite and non-negative real number.
A distance value of zero defines an identity among points (i.e. $f(x, y) = 0 \Leftrightarrow x = y$) and larger values indicate greater dissimilarity among the points.
We also require that the distance function be symmetric (i.e., $f(x, y) = f(y, x) \ \forall \ x, y \in \textbf{X}$).
In addition to these constraints, if the distance function obeys the triangle inequality (i.e. $f(x, y) \leq f(x, z) + f(z, y) \ \forall \ x, y, z \in \textbf{X}$), then it is also a \textit{distance metric}.
Similar to~\cite{yu2015entropy}, when used with distance metrics, all search algorithms in CAKES are exact.
For example, Euclidean, Levenshtein~\cite{levenshtein1966binary} and Dynamic Time Warping (DTW)~\cite{muller2007dynamic} distances are all distance metrics, while cosine distance is not a metric because it violates the triangle inequality (e.g., consider the points $x = (1, 0)$, $y = (0, 1)$ and $z = (1, 1)$ on the Cartesian plane).

The choice of an appropriate distance function varies by dataset and domain.
For example, with neural-network embeddings, one could use Euclidean (L2) or cosine distance.
With genomic or proteomic sequence data, Levenshtein, Smith-Waterman, Needleman-Wunsch or Hamming distances are useful.
With time-series data, one could use Dynamic Time Warping (DTW) or Wasserstein distance.

CAKES derives its advantages from the manifold hypothesis~\cite{fefferman2016testing}, the notion that high-dimensional data collected from constrained generating phenomena typically only occupy a low-dimensional manifold within their embedding space.
We say that such data are \textit{manifold-constrained} and have low \textit{local fractal dimension} (LFD).
In other words, we assume that the dataset is embedded in a $D$-dimensional space, but that the data only occupy a $d$-dimensional manifold, where $d \ll D$.
While we sometimes use Euclidean notions, such as voids, volumes and proximity to describe the geometric and topological properties of the clusters and manifold, CLAM and CAKES do not rely on such notions;
they serve merely as convenient and intuitive vocabulary to discuss the underlying mathematics.

@@

We stress that this concept differs from the \textit{embedding dimension} of a dataset.
To illustrate the difference, consider the SILVA 18S rRNA dataset which contains genomes with unaligned lengths of up to 3,718 base pairs and aligned length of 50,000 base pairs.
Hence, the \textit{embedding dimension} of this dataset is at least 3,718 and at most 50,000.
However, due to physical constraints (namely, biological evolution and biochemistry), the data are constrained to a lower-dimensional manifold within this embedding space.
LFD is an approximation of the dimensionality of that lower-dimensional manifold in the ``vicinity'' of a given point.
Figure~\ref{fig:results:lfd-plots} illustrates this concept on a variety of datasets, showing how real datasets uphold the manifold hypothesis.

@@

We define a \textit{cluster} as a set of points with a \textit{center} and a \textit{radius}.
The center is the geometric median of the points (or a smaller sample of the points) in the cluster, and so it is a real data point.
The radius is the maximum distance from the center to any point in the cluster.
Each non-leaf cluster has two child clusters in much the same way that a node in a binary tree has two child nodes.
Note that clusters can have overlapping volumes and, in such cases, points in the overlapping volume are assigned to exactly one of the overlapping clusters.
As a consequence, a cluster can be a proper subset of the metric ball at the same center and radius, i.e. $C(c, r) \subset B_X(c, r)$.

Hereafter, when we refer to the LFD of a cluster, it is estimated at the length scale of the cluster radius and half that radius.
This allows us to rewrite Definition~\ref{eq:methods:lfd-original} as:

@@

We start by performing a divisive hierarchical clustering on the dataset using CLAM.
The procedure is similar to that outlined in CHESS~\cite{ishaq2019clustered}, but with
better selection of poles for partitioning (see Algorithm~\ref{alg:methods:partition}) and depth-first reordering of the dataset (see Section~\ref{sec:methods:clustering:depth-first-reordering}).

@@

CLAM builds a binary tree of clusters using a divisive hierarchical clustering algorithm.

For a cluster $C$ with $|C|$ points, we begin by taking a random sample of $\sqrt{|C|}$ of its points, and computing pairwise distances between all points in this sample.
Using these distances, we compute the \textit{geometric median} of this sample; in other words, we find the point which minimizes the sum of distances to all other points in the sample.
This geometric median is the center of $C$.

@@

With no additional memory cost nor time cost for retrieving points during search, depth-first reordering offers the same time complexity as CHESS with $\mathcal{O}(n)$ space.

@@

% \begin{figure}[ht!]
%     \centering
%     \includegraphics[scale=0.3]{images/geometry/deltas.pdf}
%     \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
%     \label{fig:methods:deltas}
% \end{figure}

@@

\begin{figure}[ht!]
    \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[scale=0.4]{images/geometry/deltas.pdf}
        \caption{{\color{blue}$\delta$}, {\color{red}$\delta^{+}$}, and {\color{green}$\delta^{-}$} for a cluster $C$ and a query $q$.}
        \label{fig:methods:deltas}
      \end{subfigure}%
      \begin{subfigure}{.5\textwidth}
        \centering
        \includegraphics[scale=0.4]{images/geometry/overlapping-children-3.pdf}
        \caption{The geometry of a query ball overlapping with a cluster and either one or both of its children. Here, $l$ is the left pole, $r$ is the right pole, and $q$ is the query. Other points and distances are described in the text.}
        \label{fig:methods:overlapping-children}
      \end{subfigure}%
    \caption{CAKES uses geometric properties of clusters.}
\end{figure}

@@

From Algorithm~\ref{alg:methods:rnn-search}, we have that $H = B_X(q, \rho)$, and so $\rho$-NN search performance scales linearly with the size of the output set and exponentially with the LFD around the query.
While one might worry that the exponential scaling factor would dominate, the manifold hypothesis suggests that the LFD of real-world datasets is typically very low, and so this algorithm actually scales sub-linearly with the cardinality of the dataset.

@@

We perform some simple auto-tuning to select the optimal $k$-NN algorithm to use with a given dataset.
We start by taking the center of every cluster at a low depth, (e.g. 10), in the cluster tree as a query.
This gives us a small, representative sample of the dataset.
Using these clusters' centers as queries, and a user-specified value of $k$, we record the time taken for $k$-NN search on the sample using each of the three algorithms described in Section~\ref{sec:methods:knn-search}.
We select the fastest algorithm over all the queries as the optimal algorithm for that dataset and value of $k$.
Note that even though we select the optimal algorithm based on use with some user-specified value of $k$, we still allow search with any value of $k$.

### 2024-04-29 sections/3-datasets.tex
careful edit

To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal DNA dataset~\cite{10.1093/nar/gks1219}.
This dataset contains ribosomal DNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
Since this dataset contains many \textit{redundant}, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition \textit{criterion} in Algorithm~\ref{alg:methods:partition}.

@@

To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset~\cite{oshea2018radioml}.
The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it is comprised of 24 modulation modes at 26 different SNR levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.

>>>>

To showcase the use of CAKES with a more exotic distance function, we also benchmarked on the SILVA 18S ribosomal RNA dataset~\cite{10.1093/nar/gks1219}.
This dataset contains ribosomal RNA sequences of 2,224,640 genomes with an aligned length of 50,000 letters.
We held out a set of 1,000 random sequences from the dataset to use as queries for benchmarking.
We use Levenshtein~\cite{levenshtein1966binary} distance on the unaligned sequences to build the tree and to perform $k$-NN search.
Since this dataset contains many redundant, i.e. nearly identical, sequences, we used a maximum tree-depth of 128 as a partition criterion in Algorithm~\ref{alg:methods:partition}.

@@

To provide another example of use of CAKES with a more exotic distance function, we benchmarked on the RadioML dataset~\cite{oshea2018radioml}.
The RadioML dataset contains samples of synthetically generated signal captures of different modulation modes over a range of SNR levels.
Specifically, it comprises 24 modulation modes at 26 different signal-to-noise ratio (SNR) levels ranging from -20 dB to 30 dB, with 4,096 samples at each combination of modulation mode and SNR level.
Thus, it contains $24 \cdot 26 \cdot 4096 = 2,555,504$ samples in total.
Each sample is a 1,024-dimensional complex-valued vector, representing a signal capture (a time-series of complex-valued numbers).
We used a subset of this dataset, containing 97,920 samples at 10dB SNR and used the other 8,576 samples at 10dB SNR as a hold-out set of queries.
We use Dynamic Time Warping~\cite{muller2007dynamic} as the distance function for this dataset.

### 2024-04-29 sections/4-results.tex
careful edit

Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
Glove-25 under Cosine distance, and
Sift under Euclidean distance.

Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.

The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
The vertical axis denotes throughput in queries per second.
Both axes are on a logarithmic scale.
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
For HNSW and ANNOY, we report the recall for each measurement in the plots. For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.

@@

Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset. For Fashion-Mnist, CAKES begins exhibiting sub-linear performance starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively. 
Which one of the three CAKES algorithms is fastest also differs by dataset.
For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve. 
Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset. 
With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling. 
We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink). 
For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$. 
On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift. 
In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the Cosine distance function is not a metric).

@@

For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS do not support neither the required distance functions nor, in the case of RadioML, complex-valued data.
Due to the massive sizes of these datasets, we took random sub-samples of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance for all cardinalities.
In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.

@@

    \caption{QPS (queries per second) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.

>>>>

Figures~\ref{fig:results:fashion-mnist-scaling},~\ref{fig:results:glove-25-scaling},~and~\ref{fig:results:sift-scaling} show the scaling behavior of CAKES algorithms and existing algorithms on augmented versions of Fashion-Mnist under Euclidean distance,
Glove-25 under cosine distance, and
Sift under Euclidean distance.

Figure~\ref{fig:results:random-scaling} shows the scaling behavior of CAKES algorithms and existing algorithms on a completely randomly generated dataset with the same cardinality and dimensionality as Sift (1 million points in 128 dimensions).
Figures~\ref{fig:results:silva-scaling} and ~\ref{fig:results:radioml-scaling} show the scaling behavior of CAKES on Silva and RadioML respectively.
For these datasets, we took random sub-samples of the full datasets instead of augmenting them to higher cardinalities.

The horizontal axis in each figure shows the cardinality of the dataset augmented with synthetic points (see Section \ref{sec:methods:synthetic-data}).
The left-most point on each line is at the cardinality of the original dataset without any synthetic augmentation.
The vertical axis denotes throughput in queries per second (higher is better).
Both axes are on a logarithmic scale.
In this section, we report results only for $k$-NN search with $k = 10$, but similar plots for $k = 100$ can be found in the Supplement.
For HNSW and ANNOY, we report the recall for each measurement in the plots. For algorithms which exhibit recall greater than $0.9995$, we do not report the recall in the plots, but we do report it in the tables below.

@@

Figure~\ref{fig:results:scaling-plots} shows that for Fashion-Mnist, Glove-25, and Sift, as cardinality increases, the CAKES algorithms (Depth-First Sieve in blue, Repeated $\rho$-NN in green, and Breadth-First Sieve in purple) become faster than our Rust implementation of na\"{i}ve linear search (in orange).
Though we observe this trend, we note that the exact cardinality at which CAKES's algorithms overtake linear search differs by dataset. For Fashion-Mnist, CAKES begins outperforming linear search starting at a cardinality near $10^5$, while for Glove-25 and Sift, this happens near $10^6$ and $10^7$ respectively. 
Which one of the three CAKES algorithms is fastest also differs by dataset.
For Fashion-Mnist, Depth-First Sieve is consistently fastest, while for Glove-25, the fastest is Repeated $\rho$-NN, and for Sift, Breadth-First Sieve. 
Additionally, on all three datasets, Depth-First Sieve and Breadth-First Sieve appear to have performance which is \textit{constant} in the cardinality of the dataset. 
With Glove-25, Repeated $\rho$-NN exhibits similar constant scaling. 
We also observe that on these three datasets, for nearly all cardinalities, all three of the CAKES algorithms are faster than FAISS-Flat (in brown), and that at some cardinality, CAKES's algorithms become faster than FAISS-IVF (in pink). 
For Fashion-Mnist, CAKES becomes faster than FAISS-IVF near cardinality $10^5$, whereas for Glove-25 and Sift, this happens near cardinality $10^7$. 
On all three of these datasets, HNSW (in gray) and ANNOY (in yellow) are faster than CAKES's algorithms for all cardinalities; however, CAKES exhibits perfect or near-perfect recall on each dataset, while HNSW and ANNOY exhibit much lower recall, as shown in Table~\ref{tab:datasets:summary}.
While recall for CAKES's does \emph{not} degrade with cardinality, recall for HNSW and ANNOY does degrade with cardinality. 
At a cardinality multiplier as low as eight, HNSW and ANNOY have recall of $0.525$ and $0.857$ respectively for Fashion-Mnist, 0.607 and 0.832 for Glove-25, and 0.782 and 0.686 on Sift. 
In contrast, CAKES has perfect recall on Fashion-Mnist and Sift (because the distance function used with these datasets is a metric), and near-perfect recall on Glove-25 (because the cosine distance function is not a metric).

@@

For Silva and RadioML, we benchmarked only CAKES's algorithms because HNSW, ANNOY and FAISS support neither the required distance functions nor, in the case of RadioML, complex-valued data.
Due to the massive sizes of these datasets and challenges in generating plausible augmentations, we took random sub-samples ranging up to the entirety of the dataset--rather than augmented versions of the dataset--to examine how performance scales with cardinality.
With Silva, as shown in Figure~\ref{fig:results:silva-scaling}, we observe that for all algorithms, throughput initially seems to linearly decrease as cardinality increases, but that it seems to begin levelling off near cardinality $10^5$.
Until cardinality near $10^4$, Depth-First Sieve is the fastest CAKES algorithm, but for larger cardinalities, Repeated $\rho$-NN is the fastest CAKES algorithm.
For RadioML, as shown in Figure~\ref{fig:results:radioml-scaling}, we observe that throughput declines nearly linearly, and that the three CAKES algorithms exhibit virtually indistinguishable performance for all cardinalities.
In particular, throughput of CAKES's algorithms is identical within three significant figures for all cardinalities.

@@

    \caption{Queries per second (QPS) and Recall vs naive linear search on the Fashion-Mnist, Glove-25, Sift and Random datasets.
